Using Visual Analytics to Explain\\[1ex] Black-Box Machine Learning

### 2018-04-06 prospector/related_work.tex
typo fixing

\subsection{Algorithm Specific Model Visualization}
In the past, research has primarily focused on understanding and interacting with specific machine learning algorithms.
Often the focus is on the internal weights of the trained models.
For Bayesian networks, showing probabilities of the nodes (Becker~\etal~\cite{Becker:2001:VSB:383784.383809}) and how the input is propagated (Correa~\etal~\cite{DBLP:journals/isci/MartinsONHC13})
has been used.
For Support Vector Machines,
projection techniques (Caragea~\etal~\cite{Caragea2001}) and
Nomograms (Jakulin~\etal~\cite{Jakulin:2005:NVS:1081870.1081886})
to see the ``cut" in the data points were utilized.
Visualizing and interpreting the graph of a neural network has also been used by Tzeng and Ma~\cite{Tzeng:2005:OTB}.
Kim~\etal~\cite{kim2014bayesian,kim2015MGM} introduce graphical models that allow for
interpretability of its features.
\cite{kim2015MGM} use a colored matrix of features by category to show distinguishable features computed by their model.
Caruana~\etal~\cite{Caruana:2015:IMH:2783258.2788613} use high-performance generalized additive models (GA$^2$Ms) that allows visual inspection of the influence of its input features on the outcome much like partial dependence.

@@

Having access to the internals of a machine learning algorithm also allows direct interaction with the models and to improve them on the fly.
BaobabView (van den Elzen and van Wijk~\cite{van2011baobabview}) offers interactive decision tree construction.
Steed~\etal~\cite{steed2009guided,steed2009tropical}
guides regression model creation by enhancing interaction
with the input data.
EnsembleMatrix (Talbot~\etal~\cite{Talbot:2009:EIV:1518701.1518895})
allows users to combine already computed classification models to
build ensemble models.
Some recent interactive machine learning tools \cite{
amershi15,
Amershi:2011:DEE:2046396.2046416,
Amershi:2012:RIM:2207676.2207680,
Amershi:2011:CHF:1978942.1978966,
Kapoor:2010:IOS:1753326.1753529,
kim2015scalable,
Leung2014710,
Mishra:2015:SAI:2700171.2791022,
export:141330})
are more algorithm agnostic, but depend on general
performance measures like confusion matrices,
area under ROC curve (AUC) measures,
result distribution of data points, and
feature weights according to model independent statistics.

@@

\subsection{Probing Models}
Partial dependence was proposed by Friedman~\cite{friedman2001} for analyzing
gradient boosting models and has since been used for other models as well (\eg Ehrlinger~\cite{ehrlinger2015} uses partial dependence to analyze the behavior of random forests
in the \emph{R} package \emph{ggRandomForests}).

Cortez and Embrechts~\cite{cortez2011opening,Cortez20131} and Kim~\etal~\cite{kim2014bayesian}
use sensitivity analysis to analyze and compare machine learning models of different algorithms.
Sensitivity analysis is similar to partial dependence except that it uses a few
base vectors (usually the mean, median, or quartiles of all observed values)
instead of computing the probabilities over all input points.
This method is faster than partial dependence but may miss critical details
about the prediction function especially if the function is strongly non-linear.

Goldstein~\etal~\cite{goldstein14} extends the idea of partial dependence by using
Individual Conditional Expectation (ICE) plots which show one line for each row of the input data.
We found, however, that this often clutters the plots too much and makes them harder to interpret.
We experimented further with showing standard deviations and quartiles of the partial dependence line
but discarded this approach since the spread of the partial dependence results is always expected to
be large unless one feature dominates the classification significantly and is able to solely change
the classification even for the furthest data points.

By not restricting ourselves to sampling only the observed input space, our approach on partial dependence
enables a deeper analysis of the machine learning model. Furthermore, accepting the costs of computing
partial dependence over all input points yields proper results even for highly non-linear models,
while also not overwhelming users with too much detail.
This is strengthened even further by our novel approach of using implementation details of the inspected
models to improve the sampling and the representation of the results.

>>>>

\subsection{Algorithm Specific Model Visualization}
In the past, research has primarily focused on understanding and interacting with specific machine learning algorithms.
Often the focus is on the internal weights of the trained models.
For Bayesian networks, showing probabilities of the nodes (Becker~\etal~\cite{Becker:2001:VSB:383784.383809}) and how the input is propagated (Correa~\etal~\cite{DBLP:journals/isci/MartinsONHC13})
has been used.
For Support Vector Machines,
projection techniques (Caragea~\etal~\cite{Caragea2001}) and
Nomograms (Jakulin~\etal~\cite{Jakulin:2005:NVS:1081870.1081886})
to see the ``cut" in the instances were utilized.
Visualizing and interpreting the graph of a neural network has also been used by Tzeng and Ma~\cite{Tzeng:2005:OTB}.
Kim~\etal~\cite{kim2014bayesian,kim2015MGM} introduce graphical models that allow for
interpretability of its features.
\cite{kim2015MGM} use a colored matrix of features by category to show distinguishable features computed by their model.
Caruana~\etal~\cite{Caruana:2015:IMH:2783258.2788613} use high-performance generalized additive models (GA$^2$Ms) that allows visual inspection of the influence of its input features on the outcome much like partial dependence.

@@

Having access to the internals of a machine learning algorithm also allows direct interaction with the models and to improve them on the fly.
BaobabView (van den Elzen and van Wijk~\cite{van2011baobabview}) offers interactive decision tree construction.
Steed~\etal~\cite{steed2009guided,steed2009tropical}
guides regression model creation by enhancing interaction
with the input data.
EnsembleMatrix (Talbot~\etal~\cite{Talbot:2009:EIV:1518701.1518895})
allows users to combine already computed classification models to
build ensemble models.
Some recent interactive machine learning tools \cite{
amershi15,
Amershi:2011:DEE:2046396.2046416,
Amershi:2012:RIM:2207676.2207680,
Amershi:2011:CHF:1978942.1978966,
Kapoor:2010:IOS:1753326.1753529,
kim2015scalable,
Leung2014710,
Mishra:2015:SAI:2700171.2791022,
export:141330})
are more algorithm agnostic, but depend on general
performance measures like confusion matrices,
area under ROC curve (AUC) measures,
result distribution of instances, and
feature weights according to model independent statistics.

@@

\subsection{Probing Models}
Partial dependence was proposed by Friedman~\cite{friedman2001} for analyzing
gradient boosting models and has since been used for other models as well (\eg, Ehrlinger~\cite{ehrlinger2015} uses partial dependence to analyze the behavior of random forests
in the \emph{R} package \emph{ggRandomForests}).

Cortez and Embrechts~\cite{cortez2011opening,Cortez20131} and Kim~\etal~\cite{kim2014bayesian}
use sensitivity analysis to analyze and compare machine learning models of different algorithms.
Sensitivity analysis is similar to partial dependence except that it uses a few
base vectors (usually the mean, median, or quartiles of all observed values)
instead of computing the probabilities over all instances.
This method is faster than partial dependence but may miss critical details
about the prediction function especially if the function is strongly non-linear.

Goldstein~\etal~\cite{goldstein14} extends the idea of partial dependence by using
Individual Conditional Expectation (ICE) plots which show one line for each row of the input data.
We found, however, that this often clutters the plots too much and makes them harder to interpret.
We experimented further with showing standard deviations and quartiles of the partial dependence line
but discarded this approach since the spread of the partial dependence results is always expected to
be large unless one feature dominates the classification significantly and is able to solely change
the classification even for the furthest instance.

By not restricting ourselves to sampling only the observed input space, our approach on partial dependence
enables a deeper analysis of the machine learning model. Furthermore, accepting the costs of computing
partial dependence over all instances yields proper results even for highly non-linear models,
while also not overwhelming users with too much detail.
This is strengthened even further by our novel approach of using implementation details of the inspected
models to improve the sampling and the representation of the results.

### 2018-04-06 prospector/system.tex
typo fixing

The patient selection tab allows users to find patients they may want to inspect, based on their ground truth (\eg, whether they actually had diabetes) and their predicted risk (\eg, assessed likelihood by the predictive model of having the disease).  \prospector provides a visual summary of the patient population by providing a patient selection visualization. The visualization, as seen in Figure~\ref{figs:select}, consists of two columns dividing the population according to their ground truth
(case patients being those actually diagnosed with diabetes and control patients being not).
Each column is then separated into bins of predicted probabilities in steps of $0.1$
which can be clicked to select the group of patients fitting those criteria.  For instance, if a case patient was predicted with a low risk score, that patient would appear in the top of the right column.  If a bin is too small to provide a clickable area, a box at the side of the column is displayed to allow choosing even small populations.  The selected population is then shown next to the visualization with the individual prediction results shown for each entry.

@@

Users can change the sort order of the partial dependence bars by using the buttons at the top.  In addition to sorting by the feature weight and relevance as determined by the predictive model, users can also sort according to our local feature importance and impactful changes as described above.
If impactful changes are chosen as the sort order, the suggested changes to each feature are indicated with a white circular label in the partial dependence bar, as shown in the bottom left of Figure~\ref{figs:ui}.

>>>>

The patient selection tab allows users to find patients they may want to inspect, based on their ground truth (\eg, whether they actually had Diabetes) and their predicted risk (\eg, assessed likelihood by the predictive model of having the disease).  \prospector provides a visual summary of the patient population by providing a patient selection visualization. The visualization, as seen in Figure~\ref{figs:select}, consists of two columns dividing the population according to their ground truth
(case patients being those actually diagnosed with Diabetes and control patients being not).
Each column is then separated into bins of predicted probabilities in steps of $0.1$
which can be clicked to select the group of patients fitting those criteria.  For instance, if a case patient was predicted with a low risk score, that patient would appear in the top of the right column.  If a bin is too small to provide a clickable area, a box at the side of the column is displayed to allow choosing even small populations.  The selected population is then shown next to the visualization with the individual prediction results shown for each entry.

@@

Users can change the order of the partial dependence bars by using the buttons at the top.  In addition to sorting by the feature weight and relevance as determined by the predictive model, users can also sort according to our local feature importance and impactful changes as described above.
If impactful changes are chosen as the order, the suggested changes to each feature are indicated with a white circular label in the partial dependence bar, shown on the bottom left of Figure~\ref{figs:ui}.

### 2018-04-06 prospector/ui.tex
fix figures and some citations
typo fixing

\begin{figure}[t]
\centering
\includegraphics[width=0.8\linewidth]{prospector/ui_inspect} % 0.775
\\
\vspace*{0.75em} % 2
\includegraphics[width=0.3\linewidth]{prospector/ui_sugg} % 0.3
\hspace*{0.2\linewidth} % 18
\includegraphics[width=0.3\linewidth]{prospector/ui_change} % 0.3
% \vspace*{1em}
\caption{
The user interface of \prospector is shown at the top.
The bottom left shows suggestions on what changes (white circles) would decrease the
predicted risk the most.
The bottom right shows how the color plots change due to changing a value (namely changing the
bmi value from 0 to 1).
Fully white circles show the original value of the given patient.
}
\label{figs:ui}
\end{figure}

>>>>

\begin{figure}
\centering
\includegraphics[width=0.8\linewidth]{prospector/ui_inspect} % 0.775
\\
\vspace*{0.75em} % 2
\includegraphics[width=0.3\linewidth]{prospector/ui_sugg} % 0.3
\hspace*{0.2\linewidth} % 18
\includegraphics[width=0.3\linewidth]{prospector/ui_change} % 0.3
% \vspace*{1em}
\caption[The user interface of \prospector.]{
The user interface of \prospector is shown at the top.
The bottom left shows suggestions on what changes (white circles) would decrease the
predicted risk the most.
The bottom right shows how the color plots change due to changing a value (namely changing the
BMI value from 0 to 1).
Fully white circles show the original value of the given patient.
}
\label{figs:ui}
\end{figure}

### 2018-04-06 tex/introduction/contribution.tex
typo fixing

\subsection{\emph{\infuse}}
In the first stage we analyze and compare different feature selection strategies. Feature selection is a pre-processing step that, unlike feature engineering, algorithmically determines the possible predictive impact of a feature. Only the most impactful features are then used as input for the predictive modeling process. This is often necessary as a larger number of features require significantly more training examples to accurately represent the valid input space without losing predictive qualities. As feature selection algorithms use similar statistical tools as many predictive modeling techniques it seems plausible to infer the importance of features in the context of decision making through feature selection in a model agnostic way.

@@

\subsection{\emph{\prospector}}
In the second stage we explore the influence of features in the decision making process of predictive models through a technique called partial dependence. Partial dependence computes feature impact on model outcomes by systematically probing the model with artificial inputs. That is, while keeping the rest of the input values the same, the inspected feature assumes all possible values showing the relationship of feature values to the prediction score. Aggregated over all observed instances, the general behavior of the model with respect to a given feature can be inferred. By computing a novel feature importance score from partial dependence relationships unusual and interesting relations can be found quickly without having to inspect all, often several thousand, features.

Analyzing partial dependency relationships on diabetes prediction models trained on electronic medical records containing patient demographics, diagnoses, medications, and laboratory results, using our method we gained several insights. Firstly, we showed that logistic regression models are not expressive enough to accurately model the complex relationships of certain features to the outcome. Secondly, imputation of missing values, using the population average in laboratory results, caused the model to become uncertain close to the normal value range of laboratory tests. Lastly, we found that the model used a proxy variable indicating the number of doctor visits as predictor of the healthiness of a patient.
However, this is \emph{not} a valid predictor.
Even though a patient is likely to visit a doctor more often if she is sick, the reverse cannot be said. The above findings show how \textbf{partial dependence} with the proposed \textbf{feature importance score} allow analysts to effectively \textbf{detect model errors} related to \textbf{over-} and \textbf{under-fitting}, \textbf{imputation}, and \textbf{incorrect cause-effect relationships}.
However, a major drawback of using partial dependency is its limitations on extending it to relations of multiple features to the predicted outcome.

@@

\subsection{Model Diagnostic Workflow}
In the third stage, we overcome this limitation by leveraging instance-level explanations. Instance-level explanations are defined as the smallest change, to an instance vector, necessary to change the predicted outcome label. Aggregating over those explanations and statistically analyzing the resulting instance subsets, proves as powerful novel approach for understanding the behavior of a model. Additionally, we propose a model diagnostic workflow (see \figref{figs:workflow_intro}) that helps identify flaws in the \emph{input data}, used to train and test the model.

\begin{table}[t]
     \begin{tabular}{l|c|c|c} 
     \textbf{Approach} & \makebox[0pt][l]{\textbf{Instances}}\phantom{Aggregated} & \textbf{Aggregated} & \makebox[0pt][l]{\textbf{Global}}\phantom{Aggregated} \\ 
     \hline
     \hline
     \infuse & & & X \Tstrut\\
     \hline
     \prospector & X & & X \Tstrut\\
     \hline
     \textit{Model diagnostic workflow} & & X & \Tstrut\\
    \end{tabular}
    \centering
    \caption[Locality of decision analyses of the presented approaches.]{Locality of decision analyses of the presented approaches. \textbf{Instances} refers to analyzing decisions for one instance at a time. \textbf{Aggregated} refers to analyzing decisions for groups of instances and \textbf{Global} refers to analyzing decisions globally without inspecting decisions for individual instances.}
    \label{tab:locality}
\end{table}

We use this workflow to analyze a predictive modeling problem revolving patient admission to a hospital, of patients in the emergency department of said hospital. Accurately predicting whether a patient eventually gets admitted to the hospital helps reducing costs. The major limiting factor of this prediction task is the need for input features readily, and electronically, available at the earliest point possible. To this extend we initially used features of prescribed medications as this information is immediately available electronically. Our analysis showed that there are clear groups of patients where the predictive model is very helpful. However, there are some groups of patients where it is impossible for the predictive model to make accurate predictions given the provided information. One of those groups is patients receiving Diatrizoate Meglumine, a contrast medium for CAT or PET scans. This medication only indicates that a scan was performed, but does not carry information about the result of the scan. However, the result of the scan is the deciding factor whether a patient needs further care or can get sent home. Thus, with the given information it is impossible for the predictive model to make a decision that performs better than random guessing. Including more information in the form of additional features does indeed help with this problem but pushes the time when a decision can be made further back. One possible solution is to utilize the predictive model only for the confident groups of patients and wait for the doctors' decisions in other cases. This example illustrates that it is not only possible to \textbf{understand decision making} of a predictive model through the \textbf{model diagnostic workflow} based on \textbf{aggregated instance-level explanations}, but also how it can be applied for \textbf{semantic validation} and \textbf{feature engineering} on the input data.

>>>>

\subsection{\emph{\infuse}}
In the first stage, we analyze and compare different feature selection strategies. Feature selection is a pre-processing step that, unlike feature engineering, algorithmically determines the possible predictive impact of a feature. Only the most impactful features are then used as input for the predictive modeling process. This is often necessary as a larger number of features require significantly more training examples to accurately represent the valid input space without losing predictive qualities. As feature selection algorithms use similar statistical tools as many predictive modeling techniques it seems plausible to infer the importance of features in the context of decision making through feature selection in a model agnostic way.

@@

\subsection{\emph{\prospector}}
In the second stage, we explore the influence of features in the decision making process of predictive models through a technique called partial dependence. Partial dependence computes feature impact on model outcomes by systematically probing the model with artificial inputs. That is, while keeping the rest of the input values the same, the inspected feature assumes all possible values showing the relationship of feature values to the prediction score. Aggregated over all observed instances, the general behavior of the model with respect to a given feature can be inferred. By computing a novel feature importance score from partial dependence relationships unusual and interesting relations can be found quickly without having to inspect all, often several thousand, features.

Analyzing partial dependency relationships on diabetes prediction models trained on electronic medical records containing patient demographics, diagnoses, medications, and laboratory results, using our method we gained several insights. Firstly, we showed that logistic regression models are not expressive enough to accurately model the complex relationships of certain features to the outcome. Secondly, imputation of missing values, using the population average in laboratory results, caused the model to become uncertain close to the normal value range of laboratory tests. Lastly, we found that the model used a proxy variable indicating the number of doctor visits as predictor of the healthiness of a patient.
However, this is \emph{not} a valid predictor.
Even though a patient is likely to visit a doctor more often if she is sick, the reverse cannot be said. The above findings show how \textbf{partial dependence} with the proposed \textbf{feature importance score} allow analysts to effectively \textbf{detect model errors} related to \textbf{over-} and \textbf{under-fitting}, \textbf{imputation}, and \textbf{leaking labels} caused by incorrect cause-effect relationships.
However, a major drawback of using partial dependency is its limitations on extending it to relations of multiple features to the predicted outcome.

@@

\subsection{Model Diagnostic Workflow}
In the third stage, we overcome those limitations by leveraging instance-level explanations. Instance-level explanations are defined as the smallest change, to an instance vector, necessary to change the predicted outcome label. Aggregating over those explanations and statistically analyzing the resulting instance subsets, proves as powerful novel approach for understanding the behavior of a model. Additionally, we propose a Model Diagnostic workflow (see \figref{figs:workflow_intro}) that helps identify flaws in the \emph{input data}, used to train and test the model.

\begin{table}[t]
     \begin{tabular}{l|c|c|c} 
     \textbf{Approach} & \makebox[0pt][l]{\textbf{Instances}}\phantom{Aggregated} & \textbf{Aggregated} & \makebox[0pt][l]{\textbf{Global}}\phantom{Aggregated} \\ 
     \hline
     \hline
     \infuse & & & X \Tstrut\\
     \hline
     \prospector & X & & X \Tstrut\\
     \hline
     \textit{Model Diagnostic workflow} & & X & \Tstrut\\
    \end{tabular}
    \centering
    \caption[Locality of decision analyses of the presented approaches.]{Locality of decision analyses of the presented approaches. \textbf{Instances} refers to analyzing decisions for one instance at a time. \textbf{Aggregated} refers to analyzing decisions for groups of instances and \textbf{Global} refers to analyzing decisions globally without inspecting decisions for individual instances.}
    \label{tab:locality}
\end{table}

We use this workflow to analyze a predictive modeling problem revolving around patient admission to a hospital, of patients in the emergency department of said hospital. Accurately predicting whether a patient eventually gets admitted to the hospital helps reducing costs. The major limiting factor of this prediction task is the need for input features readily, and electronically, available at the earliest point possible. To this extend we initially used features of prescribed medications as this information is immediately available electronically. Our analysis showed that there are clear groups of patients where the predictive model is very helpful. However, there are some groups of patients where it is impossible for the predictive model to make accurate predictions given the provided information. One of those groups is patients receiving Diatrizoate Meglumine, a contrast medium for CAT or PET scans. This medication only indicates that a scan was performed, but does not carry information about the result of the scan. However, the result of the scan is the deciding factor whether a patient needs further care or can get sent home. Thus, with the given information it is impossible for the predictive model to make a decision that performs better than random guessing. Including more information in the form of additional features does indeed help with this problem but pushes the time when a decision can be made further back. One possible solution is to utilize the predictive model only for the confident groups of patients and wait for the doctors' decisions in other cases. This example illustrates that it is not only possible to \textbf{understand decision making} of a predictive model through the \textbf{Model Diagnostic workflow} based on \textbf{aggregated instance-level explanations}, but also how it can be applied for \textbf{semantic validation} and \textbf{feature engineering} on the input data.

### 2018-04-06 tex/introduction/machine_learning.tex
typo fixing

All of those measures can be read from a confusion matrix, as shown in Figure~\ref{figs:cm_intro}

@@

\begin{figure}
\centering
\includegraphics[height=10em]{figs/roc}
\caption[Example of a ROC curve.]{
Example of a ROC curve. The plot indicates the correctness of the model with the false-positive rate as horizontal axis and the true-positive rate as vertical axis. The line traces those values for different prediction thresholds. The threshold minimizing incorrect predictions is highlighted by the colored axes.
}
\label{figs:roc_intro}
\end{figure}

@@

\subsubsection{Logistic Regression}
Logistic Regression \cite{mlbook} is one of the most popular algorithms in machine learning.
This is due to its simplicity and its interpretability of the results.
Suppose $x_i$ is the value of the feature $i$ of our data, a binary classification using Logistic Regression is equivalent to:
\[
l(x) = b + \sum_{i \in \mathcal{C}} w_i x_i
\]
where $\mathcal{C}$ is the set of features, $b$ and $w_i$ are the learned weights of the model, and the outcome is determined by whether the weighted sum $l(x)$ is larger or smaller than zero.
In order to work with probabilities, have greater precision around the boundary, and to penalize points close to the boundary, a logistic function is used:
\[
P(Y=1 | x) = L(x) = \frac{1}{1 + e^{-l(x)}}
\]
which gives the likelihood of one of the classes.
In order to train the model, \ie, finding the best values of $b$ and $w_f$, the expression $P(Y=y | x)$, where $y$ represents the ground truth 0 or 1 of the training instances, needs to be maximized.
In practice, instead of maximizing $P(Y=y|x)$, the negative logarithm of this expression is minimized.
When optimized over the entire training data, this expression is called \emph{negative log likelihood} and it can be optimized via techniques such as gradient descent, to find optimal values of $b$ and $w_f$.

@@

Training a MLP or any Neural Network works as follows.
We initialize all weights with random numbers.
%\footnote{Other initialization methods such as Xavier also exist.}
For each instance of the training data set, we first compute the outcome with the current weights.
Then we compute the gradient of the loss function, (which can be either the negative log likelihood of training data labels for classification or mean of squared error for regression tasks), with respect to each parameter.
The gradient is computed on the predicted outcome $y(x)$ (\ie, values of the final layer) and the \emph{desired} outcome $\hat{y}(x)$ (\ie, value of desired class set to one and zero otherwise).
Then, going in reverse order we numerically compute this gradient for each layer and nudge the weights of those layers into the direction of the gradient.
This procedure is called back-propagation.
Ideally, we repeat back-propagation until all weights converge towards fixed numbers, at which point the model is trained.

Neural networks are a powerful machine learning tool, as they enable any function to be modeled \cite{hornik1991251} and allow for great flexibility through combining different layers.
Multi-Layer Perceptrons can potentially be used on our running Diabetes example from above, as they do not have the limitations of any of the previously discussed models.
However, neural networks typically require many thousand training examples in order to properly converge towards a performant model, making them a sub-optimal choice for tasks where acquiring data is time-consuming or expensive.

@@

Over-fitting occurs when a model learns its training data set too precisely.
For example, a Decision Tree might create one leaf node for every instance in the training data memorizing the outcomes of the training data in full.
In this case the model likely won't generalize to new, unseen data.
Thus, in order to detect over-fitting, it is common to keep a separate data set, called hold-out or test data set, at hand that is not used for training the model.
The trained model is then tested against this data set.
If the model performs well on the training data but not on the test data, then it is over-fitting.
A way to prevent this is to reduce the capacity of the model.
For example, reducing the allowed maximum depth of a Decision Tree model is a good way to reduce its capacity.

@@

\par \noindent \textbf{Leaking Labels:}
If a feature of the data is highly correlated to the outcome, a model might only learn this connection, preventing the model to generalize properly.
For example, including an additional feature ``Takes Insulin" to the Diabetes model would leak the label, as only people that are already diagnosed with Type 2 Diabetes take the drug Insulin.

\section{Visual Analytics}
\label{sec:visualanalytics}
Gaining meaningful insights about large data sets is hard.
While computing various statistics about the data is often helpful it might oversimplify or in some cases even mislead information about the data.
One example data set often used to demonstrate this is Anscombe's quartet~\cite{doi:10.1080/00031305.1973.10478966}.
The quartet consists of four different data sets that all have the same mean and sample variance for both axes, the same correlation between the axes, and the same regression lines.
However, plotting the values of the data sets reveals that each one of those data sets obeys a different, unique characteristic (see Figure~\ref{figs:anscombe}).

Visual Analytics is the area of studying interactive graphical respresentations for analyzing complex data such as Anscombe's quartet mentioned above.
This is often done in addition to and with the help of computational and statistical procedures.
Various studies \cite{doi:10.1080/01621459.1984.10478080,Treisman:1985:PPV:5088.5091,Chang:2002:GTV:820060.820062} have been performed to establish visual principles to effectively take advantage of the high processing power of the human visual system.
Additionally, interaction with the graphical representations allows to hide some of the information and reveal it through interaction to not overwhelm the analyst.
For example, ``overview first; details on demand"~\cite{Shneiderman96theeyes} is a popular mantra taking advantage of interaction.

@@

For example, with the goal of improving the accuracy of a model a human should not have the task of, \eg, manually picking the order of the features to be included in a decision tree.
A good intuition might outperform the computer momentarily but having an objective goal makes it possible to eventually find a suitable algorithm that is en par or even outperforms the human.
Additionally, manual decisions are not generalizable to other tasks and not scalable to larger datasets which in turn is the reason to use machine learning to begin with.

@@

Additionally, Visual Analytics can help humans, especially domain experts for a problem that is to be solved using machine learning, to verify the \emph{semantic accuracy} of a model.
That is, whether decisions made by the model ``make sense".
Often, the root cause of semantically inaccurate models is due to biased data.
Thus, statistical accuracy (\ie, the proportion of correctly predicted instances) of a model can be higher than of a different model, while simultaneously being semantically less accurate (see \cite{Caruana:2015:IMH:2783258.2788613,explainer}).

>>>>

All of those measures can be read from a confusion matrix, as shown in Figure~\ref{figs:cm_intro}.

@@

\begin{figure}
\centering
\includegraphics[height=10em]{figs/roc}
\caption[Example of a ROC curve.]{
Example of a ROC curve. The plot indicates the correctness of the model with the false positive rate as horizontal axis and the true positive rate as vertical axis. The line traces those values for different prediction thresholds. The threshold minimizing incorrect predictions is highlighted by the colored axes.
}
\label{figs:roc_intro}
\end{figure}

@@

\subsubsection{Logistic Regression}
Logistic Regression \cite{mlbook} is one of the most popular algorithms in machine learning.
This is due to its simplicity and its interpretability of the results.
Suppose $x_i$ is the value of the feature $i$ of our data, a binary classification using Logistic Regression is equivalent to:
\[
l(x) = b + \sum_{i \in \mathcal{C}} w_i x_i
\]
where $\mathcal{C}$ is the set of features, $b$ and $w_i$ are the learned weights of the model, and the outcome is determined by whether the weighted sum $l(x)$ is larger or smaller than zero.
In order to work with probabilities, have greater precision around the boundary, and to penalize points close to the boundary, a logistic function is used:
\[
P(Y=1 | x) = L(x) = \frac{1}{1 + e^{-l(x)}}
\]
which gives the likelihood of one of the classes.
In order to train the model, \ie, finding the best values of $b$ and $w_i$, the expression $P(Y=y | x)$, where $y$ represents the ground truth 0 or 1 of the training instances, needs to be maximized.
In practice, instead of maximizing $P(Y=y|x)$, the negative logarithm of this expression is minimized.
When optimized over the entire training data, this expression is called \emph{negative log likelihood} and it can be optimized via techniques such as gradient descent, to find optimal values of $b$ and $w_i$.

@@

Training a MLP or any Neural Network works as follows.
We initialize all weights with random numbers.
%\footnote{Other initialization methods such as Xavier also exist.}
For each instance of the training data set, we first compute the outcome with the current weights.
Then we compute the gradient of the loss function, (which can be either the negative log likelihood of training data labels for classification or the mean of squared errors for regression tasks), with respect to each parameter.
The gradient is computed on the predicted outcome $y(x)$ (\ie, values of the final layer) and the \emph{desired} outcome $\hat{y}(x)$ (\ie, value of desired class set to one and zero otherwise).
Then, going in reverse order we numerically compute this gradient for each layer and nudge the weights of those layers into the direction of the gradient.
This procedure is called back-propagation.
Ideally, we repeat back-propagation until all weights converge towards fixed numbers, at which point the model is trained.

Neural networks are a powerful machine learning tool, as they enable any function to be modeled \cite{hornik1991251} and allow for great flexibility through combining different layers.
Multi-Layer Perceptrons can potentially be used on our running Diabetes example from above, as they do not have the limitations of any of the previously discussed models.
However, Neural Networks typically require many thousand training examples in order to properly converge towards a performant model, making them a sub-optimal choice for tasks where acquiring data is time-consuming or expensive.

@@

Over-fitting occurs when a model learns its training data set too precisely.
For example, a Decision Tree might create one leaf node for every instance in the training data memorizing the outcomes of the training data in full.
In this case the model likely won't generalize to new, unseen data.
Thus, in order to detect over-fitting, it is common to keep a separate data set, called hold-out, validation, or test data set, at hand that is not used for training the model.
The trained model is then tested against this data set.
If the model performs well on the training data but not on the test data, then the model is over-fitting.
A way to prevent this is to reduce the capacity of the model.
For example, reducing the allowed maximum depth of a Decision Tree model is a good way to reduce its capacity.

@@

\par \noindent \textbf{Leaking Labels:}
If a feature of the data is highly correlated to the outcome, a model might only learn this connection, preventing the model from generalizing properly.
For example, including an additional feature ``Takes Insulin" to the Diabetes model would leak the label, as only people that are already diagnosed with Type 2 Diabetes take the drug Insulin.

\section{Visual Analytics}
\label{sec:visualanalytics}
Gaining meaningful insights about large data sets is hard.
While computing various statistics about the data is often helpful, it might oversimplify or in some cases even mislead information about the data.
One example data set often used to demonstrate this is Anscombe's quartet~\cite{doi:10.1080/00031305.1973.10478966}.
The quartet consists of four different data sets that all have the same mean and sample variance for both axes, the same correlation between the axes, and the same regression lines.
However, plotting the values of the data sets reveals that each one of those data sets obeys a different, unique characteristic (see Figure~\ref{figs:anscombe}).

Visual Analytics is the area of studying interactive graphical respresentations for analyzing complex data such as Anscombe's quartet mentioned above.
This is often done in addition to and with the help of computational and statistical procedures.
Various studies (including but not limited to \cite{doi:10.1080/01621459.1984.10478080,Treisman:1985:PPV:5088.5091,Chang:2002:GTV:820060.820062}) have been performed to establish visual principles to effectively take advantage of the high processing power of the human visual system.
Additionally, interaction with the graphical representations allows to hide some of the information and reveal it through interaction to not overwhelm the analyst.
For example, ``overview first; details on demand"~\cite{Shneiderman96theeyes} is a popular mantra taking advantage of interaction.

@@

For example, with the goal of improving the accuracy of a model a human should not have the task of, \eg, manually picking the order of the features to be included in a decision tree.
A good intuition might outperform the computer momentarily but having an objective goal makes it possible to eventually find a suitable algorithm that is en par or even outperforms the human.
Additionally, manual decisions are not generalizable to other tasks and not scalable to larger data sets which in turn is the reason to use machine learning to begin with.

@@

Additionally, Visual Analytics can help humans, especially domain experts with a problem that is to be solved using machine learning, to verify the \emph{semantic accuracy} of a model.
That is, whether decisions made by the model ``make sense".
Often, the root cause of semantically inaccurate models is due to biased data.
Thus, statistical accuracy (\ie, the proportion of correctly predicted instances) of a model can be higher than of a different model, while simultaneously being semantically less accurate (see \cite{Caruana:2015:IMH:2783258.2788613,explainer}).
